Sketch Recognition and Understanding

### 2009-08-18 trunk/draft/System.tex
After all edit

 As shown in Figure \ref{fig:Blockdiagram} first an attempt is made to fit the stroke points into a curve or an ellipse using a minimum square error fitting algorithm \cite{chernov-2003}. If the stroke proved to be an elliptical arc then the segmentation process ends and the system proceeds to the next step. Otherwise, the stroke is passed to two particle swarm algorithms that divide the stroke to either a polygon or a set of lines and curves. The algorithms takes the stroke points along with the possible dominant points $P_{pd}$ computed during preprocessing then produce a set of dominant points which are connected by either lines or curves. The two algorithms will generate two segmentations, the system will choose the segmentation that has the minimum segmentation error.%
% First an attempts the segmentation process is divided into two steps ellipse fitting and curve segmentation. In the first step, the ellipse fitting process tries to fit the stroke into an ellipse. If the system fails it passes the stroke to the second step; curve segmentation which consists of two PSO segmentation algorithms. The two algorithms will generate two segmentations, the system will choose the segmentation that has the minimum segmentation error.%  generate the segmentations,  the minimum error will be the chosen segmentation. % is the ellipse fitting

@@

\section{Possible Dominant Point Extraction}
\label{sec:Preprocessing}
%Some data is extracted from the points in the stroke.
The preprocessing step captures the points from the pointing devices then compute the information needed to determine a set of  \textit{Possible Dominant Points $P_{pd}$} which guides the segmentation step. It is noted as a person draws a shape the pen slows down near corners and picks up speed when drawing straight lines. Therefore, the speed information is widely used to identify the shape corners and edges \cite{earlyprocess}. The curvature information is used to determine the points with high angular changes along the path of points \cite{meanshift10}. Those observations help detecting the dominant points as the points with lower speeds and high curvatures.

@@

\section{Segmentation}
\label{sec:Segmentation}
The segmentation tries to divide the stroke into a set of primitives. As shown in Figure \ref{fig:Blockdiagram} first an attempt is made to fit the stroke points into a curve or an ellipse using a minimum square error fitting algorithm \cite{chernov-2003}. If the stroke proved to be an ellipse arc then the segmentation process ends and the system proceeds to the next step. Otherwise, the stroke is passed to two particle swarm algorithms that divide the stroke to either lines or lines and curves. The algorithms take the stroke points along with the possible dominant points $P_{pd}$ computed during preprocessing then produce a set of dominant points which are connected by either lines or curves. The next section describes the ellipse detection algorithm and the two particle swarm algorithms used to divide the stroke.
%If the stroke proved to be an ellipse then the segmentation process ends and the system proceeds to the next step. Otherwise, the stroke is passed into two particle swarm algorithms that will divide the stroke to either lines or lines and curves (see the block diagram in fig. \ref{fig:Blockdiagram} ). 
%The algorithms takes the stroke points along with the possible dominant points computed then produce a set of dominant points which are connected with either lines or curves (see fig. \ref{fig:Blockdiagram}).  The next sections will describe the ellipse detection algorithm and both the two particle swarm algorithms used to divide the stroke.
%After computing the primary data the system now tries \\
%Paragraphs describe the segmentation algorithm.

@@

%The process starts by computing the center of the stroke bounding box. The bounding box center point is used as the first estimation of the ellipse center. The axes of the ellipse are estimated as $width/2$ and $height/2$ of the stroke bounding box. The least square fitting algorithm is used to minimize the fitting error of the ellipse equation \cite{chernov-2003}.  
This process tries to fit the stroke points into an ellipse arc; it starts with computing the center of the stroke bounding box. The bounding box center point is used as the first estimation of the center of the ellipse. The axes of the ellipse are estimated as the $width/2$ and $height/2$ of the stroke bounding box where $width$ is width and $height$ is height of the bonding box. The least square fitting algorithm \cite{chernov-2003} is used to minimize the fitting error of the ellipse Equation (\ref{eq:circleFit})

@@

After the best fit segmentaion is choosen the paramter of each primitive is computed. For example, if the stroke is represented as an ellispe, the center and axii of the ellipse is computed and confirmed. For lines the slope, equation, start point and end point are computed.  These values and paramter will help determine the spatial relations between different segments in the feature extraction step. 
% The segmentation which crossponds to the minimum $E_{tot}$ is the best segmentation.  
%\begin{equation}
%\label{eq:bestFit1}
%E_{area}=A_{st}-A_{sg}
% \end{equation}
% \begin{equation}
% where $A_{st}$ is the area of the input stroke, $A_{sg}$ is area of computed segmentation, $E_{alg}$ is the Segmentation Error.

@@

After the segmentation, each stroke is represented as a list of segments $L_s$. A segment can be either an ellipse, a line or a circular arc. The list of segments $L_s$ represents the segmentaion of the storke. When a user draw another stroke the generated segmentaion of the storke is appended to the list of segments.  When the user finish drawing the symbol the segment list $L_s$ will contain all segments of the drawn symbol. The relation and properties between different segments are extracted from the segment list $L_s$.  This step is considered as feature extration which is dircribed in details in the next sections.

@@

\section{Classification}%%Feature Set
\label{sec:Classification}
The system uses support vector machine (SVM) classifier with Linear kernel (Linear kernel) \cite{libsvm}. SVM is a binary classifier that maximizes the margin between the decision boundary and the support vectors. To enable SVM to learn multi classes we used one versus one strategy for combining set of binary classifier. This strategy trains $ \frac{n(n-1)}{{2}}$ binary classifier each distinguish between only two shapes. The final result is based on a number of votes each shape has from all classifiers. We used the library implementation of LibSVM \cite{libsvm} and its implementation of the one-vs-one strategy. Section \ref{Sec:SVMdetail} discribe in details the classifier, how classes (classification) are determined and the reasons behind using this specific classification methods.

\subsection {Trainig the SVM}
\subsection {Testing}

\subsection{Support Vector Machines}
\label{Sec:SVMdetail}
The foundation of Support Vector Machines (SVM) were developed by Vapnik \cite{svmintroduce} in 1995. It is gainig increase popularity due to the high performance and efficient implementation. The goal of SVM algorithms is to generate a sperating plane between two different classes which are represented using a set of n-dimentional vectors. The generated plane is used as a good classifier to the unseen exmaples. However, there can be more than one plane that can separates the data therefore SVM attempts to find the plane that maximize the margin (distance between classifier and nearest sample) between the two classes \ref{fig:svm1}. To calcuate the margin, two different hyperplanes are constructed, one plane on each side of the data, these two planes are then moved to generenate the maximum margin possilbe.  
\begin{figure}
	\centering
		\includegraphics{afterDefense/svm1.jpg.eps}
	\caption{Different Classifier planes }
	\label{fig:svm1}
\end{figure}

Let us consider  a two class classification task with datapoints $x_i$ where $i=1,\cdots,m$ having corresponding labels $y_i=\pm1$ where if  $y_i=1$ if $x_i$ belonged to first class and $y_i=-1$ if $x_i$ belonged to second class. Each data  point $x_i$ is represented in $N$ dimensional input or attribule space. Let the classification plane be represented as function. 
\begin{equation}
 y=f(x)=sign(\omega \cdot x - b)
\label{eq:planeEq}
\end{equation}

 The orientation of the speration plane is deteremined using the vector $\omega$ and the offest of the plane from the orgin is determined using the scalar $b$.   Lets make an assumption that the datatpoint are linearly seperable, i.e. there are infinite possibilities to find a plane that seperates and correctly classify the training data. This is idea is illustrated in Figure\ref{fig:svm1} where two different seperating planes can correctly classify the data. It is intuitively noted from the figure that the solid line is more likely to generalize more data. This means less classification error will be introduced. The geometrical characterization of this plane means the plane furthest from both classes.     % Which one is preferable ? Intuitively one perfers the solid pane since small perturbations of any point would not introduce misclassifications errors. Without any information, the solid plane is more likely to generalize better on future data. Geometrically we can characterize the solid plane as being furthuret from both classes.  plane supports a class if all points in that class are on one side of that plane.

The main problem of SVM is constructing the plane that is furthuest from both classes. The main approach is to use two parallel supporting planes and maximize the margin between them. All point in a class should be in one side of the plane in order to be a supporting plane.  Therefore, for the points with the class label $+1$ we would like there exist $\omega$ and $b$ such that $\omega \cdot x_i\> b$ or $\omega \cdot x_i -b \> 0$%, and for the points with class label $-1$ we would like there exist $\omega$ and $b$ such that $\omega \cdot x_i \< b $.
Suppose that smallest value of $\mid \omega \cdot x_i -b\mid$ is $\kappa$, then when $ \omega \cdot x_i -b \geq \kappa$. The argument inside the decision function is invariant under positive rescalling so we will implicity fix a scale by requiring $\omega \cdot x_i -b \geq 1 $. Similary the points with class label $-1$ requires $\omega \cdot x_i -b \leq -1 $. To find the plane furthest from both classes, we need to maximize the distance between the support planes for each class as illustrated in Figure \ref{fig:svm2}. To accomplish this, both planes are pushed apart untill they bump into small number of datapoints ( the support vecotrs ) for each class. Figure \ref{fig:svm2} shows the support vectors as filled datapoints, 
\begin{figure}
	\centering
		\includegraphics{afterDefense/svm2.jpg.eps}
	\caption{Margin of Support planes }
	\label{fig:svm2}
\end{figure}

@@

The Support vectors that are pumped into the supprting plane deteremine the margin. This is a primal form Quadratic Programming (QP) problem. The concept of duality can also used to wields the problem to a Largrnigian dual QP problem in Equation \ref{eq:dual}. 
% maximumm margin between parallel 
\begin{equation}
\begin{array}{c}
\min_{\alpha} \ \ \   \frac{1}{2} \sum_{i=1}^{m}{\sum_{j=1}^{m}{y_iy_j\alpha_i\alpha_jx_ix_j}} - \sum_{i=1}^{m}{\alpha_i} \\
s.t. \ \ \sum_{i=1}^{m}{y_i\alpha_i}=0   \\
  \alpha_i \geq 0  \ \ \ i=1,\cdots , m 
  \end{array} 
\label{eq:dual}
\end{equation}

Both the dual form and the primal form weilds to the same plane $\omega=\sum_{i=1}^{m}y_i\alpha_i x_i$ and threshold $b$ determined by the support vectors (for which $\alpha_i\geq 0$).

\subsubsection{Theoretcical foundation}

The generalization error on future points and not only in the training set can be obtained as proved by the statical learning theory \cite{BennettSVMP2000}.  The generalized error bound is a function of the misclassified error on the training data and the terms that measure the complexiy or capacity of the classification. In linear functions, using planes to maximize the margin of seperation reduces the complexiy of the functions. Thus, by explicity maximizing the margin we are in the same time  minimizing bound on the generalization error and therefore expect better generalization with high probability. The dimensionality of the data dose not affect the size of the margin which   results in good performance  for very high dimensional data (i.e, with a very large number of attribute). In a sense, this also reduces the  problems caused by over fitting of high - dimensional data.  A large volume of literature  , e.g., \cite{bookSVMoverfit11,statisticalSVM1} refers to this problem,  for more technical discussion of the   statistical theroy.

%Using geomteric arguments we can gain in sights to some of these results. Classification function that have capcity to fit the training data are more likely to overfit  resulting in poor generalization.

\subsubsection{Linearly Inseperable Case}
In all previous sections we assumed that the problem of classifiying two set is linearly separable. The assumption is not true in most casses, thus, trying to construct a plane to seperate the two dataset will fail. Figure \ref{fig:svm5} illustrates that we cannot construct a plane that can seperate the two sets correctly. The constrains for constructing the planes must be relaxed since the QP task is not feasible for the linearly inseperable case.  Consider the linear inseperable problem in Figure\ref{fig:svm5}, in the ideal case there will be no point missclassified and no point falls on the margin. But we must relax the constrain that insure that each point in on the approiate side of tis supporitng plane. Every point that falls on the wrong side of the supporting plane is considered an error. The problem is now to maximize the margin and minimize the error.

\begin{figure}
	\centering
		\includegraphics{afterDefense/svm5.jpg.eps}
	\caption{Linearly Inseperable}
	\label{fig:svm5}
\end{figure}
A minor change in the supporting plane QP problem can accomplish this. A nonegative slack or error variable $z_i$ is added to each constraint.  A weighted penalty term is also added to the objective as follows:
 \begin{equation}
 \begin{array}{c}
\min_{w,b,z}  \quad \frac{{\parallel \omega \parallel}^2}{2}+C \sum_{i=1}^{l}z_i  \\
s.t.  \quad y_i(\omega \cdot x_i -b)+z_i \geq 1  \\
    z_i \geq 0  i=1,\cdots,m
\end{array} 
\label{eq:inseperable}
\end{equation}

@@

\subsubsection {Non linear function Using Kernels}
Figure \ref{fig:svm6} shows a classification problem that cannot be seperated using linear space. A higher order plane can work better in similar casses. Converting a linear classification algorithm into non linear classification algoirhtm  can be achieved by adding addional attribute to the data that are nonlinear functions of the original data. The expanded dataset features space can be seperated using an  existing linear classification algorithm.  The  linear classication that is applied on the feautre space produce a non linear classification on the orginal inputspace. In two dimensional vector space with attributes $r$ and $s$, to construct a quadratic  discriminant  simply map the original two dimensional input space $\left[r,s\right]$ to  five dimensional features space $\left[r,s,rs,r^2,s^2\right]$ and construct a linear discriminiant in that space. specifically, define : $\vartheta(x):R^2  \rightarrow R^5$ then 
\begin{equation}
\begin{array}{l}
	x=[r,s] \\
	\omega \cdot x=\omega_1 r+ \omega_2 s \\
	\downarrow \\
	\vartheta(x)=\left[r,s,rs,r^2,s^2\right] \\

@@

For higher dimentional datasets, this method of nonlinear mapping  has two potential problems stemming from the fact that dimensionality of the feature space explones exponentially. The first problem is that overfitting becomes a problem. However, since SVM rely on margin maximization they are largely immune to this porblem. Provided an appropriate value of paramter $C$ is chosen. The second concern is that it is the computaional complexity of   computing $\vartheta(x)$ but SVM use kernels to get around this problem.

Examine what happens when athe nonlinear mapping is introduced to equation \ref{eq:dual}. Let us define : $\vartheta(x):R^2\rightarrow R^5 \quad n>> n$  we need to optimize

\begin{equation}
\begin{array}{c}
\min_{\alpha} \ \ \   \frac{1}{2} \sum_{i=1}^{m}{\sum_{j=1}^{m}{y_iy_j\alpha_i\alpha_j\vartheta(x_i)\vartheta(x_j)}} - \sum_{i=1}^{m}{\alpha_i} \\
s.t. \ \ \sum_{i=1}^{m}{y_i\alpha_i}=0   \\
 C \geq \alpha_i \geq 0  \ \ \ i=1,\cdots , m 
 \end{array} 
\label{eq:dualkernel}
\end{equation}
Notice that the mapped data only occurs as inner product in the objective. Now we apply a little mathematically rigorous magie known as hilbern schmidt kernels, first applied to SVMs in \cite{svmintroduce}.  By Mercer's theorem, we knoew that for ccertain mapping $\theta$ and any two points $u$ and $v$, the inner product of the the mapped points can be evalutated using the kernel function without ever explicity knowing ht emapping , e.g., $\theta(u)\theta(v)\equiv K(u,v)$. Some of the more popular known kernels are given below. New kernel are being developed to fit domian specific requriments. 
\begin{center}
	\begin{tabular}{cc}
	$\vartheta(u)$  & $K(u,v)$ \\ \hline 
	Degree $d$ polynomial & $\left(u\cdot v +1\right)^d$ \\ 
	Radial Basis Function & $\exp \left( - \frac{\left\|u-v\right\|^2}{2\sigma}\right) $ \\ 
		Two layer Neural Network& $ sigmoid \left(\eta\left(u\cdot v +1\right)+c\right)$ \\

	\end{tabular}
\end{center}
To change from a linear to nonlinear classifier, one must only substitute a kernel evaluation in the objective insted of the original dot product. Thus by changing kernels we can get different hightly nonlinear classifiers. No algorithmic changes are required from linear clase other than substitue of a kernel evaluation for simple dot product. All the benefits fo the original linear SVM methods are maintained. We can train a highly nonlinear classification function such as a ploynomial or a raadial basis function machine, or a sigmoidal neural network using robust, efficient algorithsm that have no problem with local minima. By using kernel substitution a linear algorithm (only capable of handling seperable data) can be turned into a general nonlinear algorithm. 
%\begin{equation}
%\begin{array}{cc}
%\underline{&}\\  
% & \left(u\cdot v +1\right)^d \\
%Radial Basis Function & \exp \left( - \frac{\left\|u-v\right\|^2}{2\sigma}\right) \\
%Two layer Neural Network & sigmoid \left(\eta\left(u\cdot v +1\right)+c\right) 
% \end{array} 
%\label{eq:diffkernel}
%\end{equation}

%The foundations of Support Vector Machines (SVM) have
\subsubsection{Classification}
The resulting SVM method can be summarized as follows 
\begin{enumerate}
	\item Select paramter C representign the tradeoff between minimizign the training set error and maximizing the margin. Select the kernel function and any kernel paramters. For example for the radial basis function kernel, one must select width of the gaussian $\gamma$. 
	\item Solve dual QP Equation \ref{} or an alterantive SVM formulation using an approriate quadratic programming or linear programming alogrithm. 
	\item Recover th primal threshold variable $b$ using the support vectors. 
	\item Classify new points using $f(x)=sign\left(\sum{i}y_i\alpha_i K\left(x,x_i\right)-b\right)$
\end{enumerate}

The paramters in step 1. are selected using cross valiation if sufficent data are avaliable. %However, recent model  selectin strategies can

>>>>

 As shown in Figure \ref{fig:Blockdiagram} first an attempt is made to fit the stroke points into a curve or an ellipse using a minimum square error fitting algorithm \cite{chernov}. If the stroke proved to be an elliptical arc then the segmentation process ends and the system proceeds to the next step. Otherwise, the stroke is passed to two particle swarm algorithms that divide the stroke to either a polygon or a set of lines and curves. The algorithms takes the stroke points along with the possible dominant points $P_{pd}$ computed during Possible Dominant Point Extraction then produce a set of dominant points which are connected by either lines or curves. The two algorithms will generate two segmentations, the system will choose the segmentation that has the minimum segmentation error.%

% First an attempts the segmentation process is divided into two steps ellipse fitting and curve segmentation. In the first step, the ellipse fitting process tries to fit the stroke into an ellipse. If the system fails it passes the stroke to the second step; curve segmentation which consists of two PSO segmentation algorithms. The two algorithms will generate two segmentations, the system will choose the segmentation that has the minimum segmentation error.%  generate the segmentations,  the minimum error will be the chosen segmentation. % is the ellipse fitting

@@

\section{Possible Dominant Point Extraction}
\label{sec:Preprocessing}
%Some data is extracted from the points in the stroke.
The Possible Dominant Point Extraction step captures the points from the pointing devices then compute the information needed to determine a set of  \textit{Possible Dominant Points $P_{pd}$} which guides the segmentation step. It is noted as a person draws a shape the pen slows down near corners and picks up speed when drawing straight lines. Therefore, the speed information is widely used to identify the shape corners and edges \cite{earlyprocess}. The curvature information is used to determine the points with high angular changes along the path of points \cite{meanshift10}. Those observations help detecting the dominant points as the points with lower speeds and high curvatures.

@@

\section{Segmentation}
\label{sec:Segmentation}
The segmentation tries to divide the stroke into a set of primitives. As shown in Figure \ref{fig:Blockdiagram} first an attempt is made to fit the stroke points into a curve or an ellipse using a minimum square error fitting algorithm \cite{chernov}. If the stroke proved to be an ellipse arc then the segmentation process ends and the system proceeds to the next step. Otherwise, the stroke is passed to two particle swarm algorithms that divide the stroke to either lines or lines and curves. The algorithms take the stroke points along with the possible dominant points $P_{pd}$ computed during Possible Dominant Point Extraction then produce a set of dominant points which are connected by either lines or curves. The next section describes the ellipse detection algorithm and the two particle swarm algorithms used to divide the stroke.
%If the stroke proved to be an ellipse then the segmentation process ends and the system proceeds to the next step. Otherwise, the stroke is passed into two particle swarm algorithms that will divide the stroke to either lines or lines and curves (see the block diagram in fig. \ref{fig:Blockdiagram} ). 
%The algorithms takes the stroke points along with the possible dominant points computed then produce a set of dominant points which are connected with either lines or curves (see fig. \ref{fig:Blockdiagram}).  The next sections will describe the ellipse detection algorithm and both the two particle swarm algorithms used to divide the stroke.
%After computing the primary data the system now tries \\
%Paragraphs describe the segmentation algorithm.

@@

%The process starts by computing the center of the stroke bounding box. The bounding box center point is used as the first estimation of the ellipse center. The axes of the ellipse are estimated as $width/2$ and $height/2$ of the stroke bounding box. The least square fitting algorithm is used to minimize the fitting error of the ellipse equation \cite{chernov-2003}.  
This process tries to fit the stroke points into an ellipse arc; it starts with computing the center of the stroke bounding box. The bounding box center point is used as the first estimation of the center of the ellipse. The axes of the ellipse are estimated as the $width/2$ and $height/2$ of the stroke bounding box where $width$ is width and $height$ is height of the bonding box. The least square fitting algorithm \cite{chernov} is used to minimize the fitting error of the ellipse Equation (\ref{eq:circleFit})

@@

After the best fit segmentation is chosen the parameter of each primitive is computed. For example, if the stroke is represented as an ellipse, the center and axii of the ellipse is computed and confirmed. For lines the slope, equation, start point and end point are computed.  These values and parameter will help determine the spatial relations between different segments in the feature extraction step. 
% The segmentation which crossponds to the minimum $E_{tot}$ is the best segmentation.  
%\begin{equation}
%\label{eq:bestFit1}
%E_{area}=A_{st}-A_{sg}
% \end{equation}
% \begin{equation}
% where $A_{st}$ is the area of the input stroke, $A_{sg}$ is area of computed segmentation, $E_{alg}$ is the Segmentation Error.

@@

After the segmentation, each stroke is represented as a list of segments $L_s$. A segment can be either an ellipse, a line or a circular arc. The list of segments $L_s$ represents the segmentation of the stroke. When a user draw another stroke the generated segmentation of the stroke is appended to the list of segments.  When the user finish drawing the symbol the segment list $L_s$ will contain all segments of the drawn symbol. The relation and properties between different segments are extracted from the segment list $L_s$.  This step is considered as feature extraction which is described in details in the next sections.

@@

\section{Classification}%%Feature Set
\label{sec:Classification}
The system uses support vector machine (SVM) classifier with Linear kernel (Linear kernel) \cite{libsvm}. SVM is a binary classifier that maximizes the margin between the decision boundary and the support vectors. To enable SVM to learn multi classes we used one versus one strategy for combining set of binary classifier. This strategy trains $ \frac{n(n-1)}{{2}}$ binary classifier each distinguish between only two shapes. The final result is based on a number of votes each shape has from all classifiers. We used the library implementation of LibSVM \cite{libsvm} and its implementation of the one-vs-one strategy. Section \ref{Sec:SVMdetail} describe in details the classifier, how classes (classification) are determined and the reasons behind using this specific classification methods.

\subsection {Training and Testing}
A training set that contains positive samples of each category in the data set is used to train the SVM classifier.  The training starts after computing the feature vector of each sample. The system trains $ \frac{n(n-1)}{{2}}$ binary classifiers each distinguish between only two categories. The goal of the training process is to generate a linear SVM plane that maximize margin between the two categories.  A cross validation procedure is implemented to choose the best value for the $C$ parameter in SVM classifier.( $C=100$ was the best value) 
=
For classification the decision is made from producing the feature vector of the unknown symbol to the trained $\frac{n(n-1)}{{2}}$ binary classifiers. The final decision is given using simple voting strategy.

\subsection{Support Vector Machines}
\label{Sec:SVMdetail}
The foundation of Support Vector Machines (SVM) were developed by Vapnik \cite{svmintroduce} in 1995. It is gaining increase popularity due to the high performance and efficient implementation. The goal of SVM algorithms is to generate a separating plane between two different classes which are represented using a set of n-dimensional vectors. The generated plane is used as a good classifier to the unseen examples. However, there can be more than one plane that can separates the data therefore SVM attempts to find the plane that maximize the margin (distance between classifier and nearest sample) between the two classes \ref{fig:svm1}. To calculate the margin, two different hyper planes are constructed, one plane on each side of the data, these two planes are then moved to generate the maximum margin possible.  
\begin{figure}
	\centering
		\includegraphics{afterDefense/svm1.jpg.eps}
	\caption{Different Classifier planes }
	\label{fig:svm1}
\end{figure}

Let us consider a two class classification task with data points $x_i$ where $i=1,\cdots,m$ having corresponding labels $y_i=\pm1$ where if  $y_i=1$ if $x_i$ belonged to first class and $y_i=-1$ if $x_i$ belonged to second class. Each data point $x_i$ is represented in $N$ dimensional input or attribute space. Let the classification plane be represented as function. 
\begin{equation}
 y=f(x)=sign(\omega \cdot x - b)
\label{eq:planeEq}
\end{equation}

 The orientation of the separation plane is determined using the vector $\omega$ and the offset of the plane from the origin is determined using the scalar $b$.   Let's make an assumption that the data point are linearly separable, i.e. there are infinite possibilities to find a plane that separates and correctly classify the training data. This is idea is illustrated in Figure\ref{fig:svm1} where two different separating planes can correctly classify the data. It is intuitively noted from the figure that the solid line is more likely to generalize more data. This means less classification error will be introduced. The geometrical characterization of this plane means the plane furthest from both classes.     % Which one is preferable ? Intuitively one prefers the solid pane since small perturbations of any point would not introduce misclassifications errors. Without any information, the solid plane is more likely to generalize better on future data. Geometrically we can characterize the solid plane as being furtherer from both classes.  Plane supports a class if all points in that class are on one side of that plane.

The main problem of SVM is constructing the plane that is furthest from both classes. The main approach is to use two parallel supporting planes and maximize the margin between them. All point in a class should be in one side of the plane in order to be a supporting plane.  Therefore, for the points with the class label $+1$ we would like there exist $\omega$ and $b$ such that $\omega \cdot x_i\> b$ or $\omega \cdot x_i -b \> 0$%, and for the points with class label $-1$ we would like there exist $\omega$ and $b$ such that $\omega \cdot x_i \< b $.
Suppose that smallest value of $\mid \omega \cdot x_i -b\mid$ is $\kappa$, then when $ \omega \cdot x_i -b \geq \kappa$. The argument inside the decision function is invariant under positive rescaling so we will implicitly fix a scale by requiring $\omega \cdot x_i -b \geq 1 $. Similarly the points with class label $-1$ requires $\omega \cdot x_i -b \leq -1 $. To find the plane furthest from both classes, we need to maximize the distance between the support planes for each class as illustrated in Figure \ref{fig:svm2}. To accomplish this, both planes are pushed apart until they bump into small number of data points ( the support vectors ) for each class. Figure \ref{fig:svm2} shows the support vectors as filled data points, 
\begin{figure}
	\centering
		\includegraphics{afterDefense/svm2.jpg.eps}
	\caption{Margin of Support planes }
	\label{fig:svm2}
\end{figure}

@@

The Support vectors that are pumped into the supporting plane determine the margin. This is a primal form Quadratic Programming (QP) problem. The concept of duality can also used to wields the problem to a Largrnigian dual QP problem in Equation \ref{eq:dual}. 
% maximum margin between parallel 
\begin{equation}
\begin{array}{c}
\min_{\alpha} \ \ \   \frac{1}{2} \sum_{i=1}^{m}{\sum_{j=1}^{m}{y_iy_j\alpha_i\alpha_jx_ix_j}} - \sum_{i=1}^{m}{\alpha_i} \\
s.t. \ \ \sum_{i=1}^{m}{y_i\alpha_i}=0   \\
  \alpha_i \geq 0  \ \ \ i=1,\cdots , m 
  \end{array} 
\label{eq:dual}
\end{equation}

Both the dual form and the primal form wields to the same plane $\omega=\sum_{i=1}^{m}y_i\alpha_i x_i$ and threshold $b$ determined by the support vectors (for which $\alpha_i\geq 0$).

\subsubsection{Theoretical foundation}

The generalization error on future points and not only in the training set can be obtained as proved by the statistical learning theory \cite{BennettSVMP2000}.  The generalized error bound is a function of the misclassified error on the training data and the terms that measure the complexity or capacity of the classification. In linear functions, using planes to maximize the margin of separation reduces the complexity of the functions. Thus, by explicitly maximizing the margin we are in the same time minimizing bound on the generalization error and therefore expect better generalization with high probability. The dimensionality of the data does not affect the size of the margin which   results in good performance for very high dimensional data (i.e, with a very large number of attribute). In a sense, this also reduces the problems caused by over fitting of high - dimensional data.  A large volume of literature, e.g., \cite{bookSVMoverfit11,statisticalSVM1} refers to this problem,  for more technical discussion of the   statistical theory.

%Using geometric arguments we can gain insights to some of these results. Classification function that have capacity to fit the training data are more likely to over fit resulting in poor generalization.

\subsubsection{Linearly Inseparable Case}
In all previous sections we assumed that the problem of classifying two set is linearly separable. The assumption is not true in most cases, thus, trying to construct a plane to separate the two dataset will fail. Figure \ref{fig:svm5} illustrates that we cannot construct a plane that can separate the two sets correctly. The constrains for constructing the planes must be relaxed since the QP task is not feasible for the linearly inseparable case.  Consider the linear inseparable problem in Figure\ref{fig:svm5}, in the ideal case there will be no point misclassified and no point falls on the margin. But we must relax the constrain that insure that each point in on the appropriate side of it is supporting plane. Every point that falls on the wrong side of the supporting plane is considered an error. The problem is now to maximize the margin and minimize the error.

\begin{figure}
	\centering
		\includegraphics{afterDefense/svm5.jpg.eps}
	\caption{Linearly Inseparable}
	\label{fig:svm5}
\end{figure}
A minor change in the supporting plane QP problem can accomplish this. A non negative slack or error variable $z_i$ is added to each constraint.  A weighted penalty term is also added to the objective as follows:
 \begin{equation}
 \begin{array}{c}
\min_{w,b,z}  \quad \frac{{\parallel \omega \parallel}^2}{2}+C \sum_{i=1}^{l}z_i  \\
s.t.  \quad y_i(\omega \cdot x_i -b)+z_i \geq 1  \\
    z_i \geq 0  i=1,\cdots,m
\end{array} 
\label{eq:inseperable}
\end{equation}

@@

\subsubsection {Non linear function Using Kernels}
Figure \ref{fig:svm6} shows a classification problem that cannot be separated using linear space. A higher order plane can work better in similar cases. Converting a linear classification algorithm into non linear classification algorithm can be achieved by adding additional attribute to the data that are nonlinear functions of the original data. The expanded dataset features space can be separated using an existing linear classification algorithm.  The linear classification that is applied on the feature space produces a non linear classification on the original input space. In two dimensional vector space with attributes $r$ and $s$, to construct a quadratic  discriminate  simply map the original two dimensional input space $\left[r,s\right]$ to  five dimensional features space $\left[r,s,rs,r^2,s^2\right]$ and construct a linear discriminate in that space. specifically, define : $\vartheta(x):R^2  \rightarrow R^5$ then 
\begin{equation}
\begin{array}{l}
	x=[r,s] \\
	\omega \cdot x=\omega_1 r+ \omega_2 s \\
	\downarrow \\
	\vartheta(x)=\left[r,s,rs,r^2,s^2\right] \\

@@

For higher dimensional datasets, this method of nonlinear mapping has two potential problems stemming from the fact that dimensionality of the feature space explodes exponentially. The first problem is that over fitting becomes a problem. However, since SVM rely on margin maximization they are largely immune to this problem. Provided an appropriate value of parameter $C$ is chosen. The second concern is that it is the computational complexity of   computing $\vartheta(x)$ but SVM use kernels to get around this problem.

Examine what happens when  the nonlinear mapping is introduced to equation \ref{eq:dual}. Let us define: $\vartheta(x):R^2\rightarrow R^5 \quad n>> n$  we need to optimize

\begin{equation}
\begin{array}{c}
\min_{\alpha} \ \ \   \frac{1}{2} \sum_{i=1}^{m}{\sum_{j=1}^{m}{y_iy_j\alpha_i\alpha_j\vartheta(x_i)\vartheta(x_j)}} - \sum_{i=1}^{m}{\alpha_i} \\
s.t. \ \ \sum_{i=1}^{m}{y_i\alpha_i}=0   \\
 C \geq \alpha_i \geq 0  \ \ \ i=1,\cdots , m 
 \end{array} 
\label{eq:dualkernel}
\end{equation}
Notice that the mapped data only occurs as inner product in the objective. Now we apply a little mathematically magic known as hilbern schmidt kernels, first applied to SVMs in \cite{svmintroduce}.  By Mercer's theorem, we know that for certain mapping $\theta$ and any two points $u$ and $v$, the inner product of the mapped points can be evaluated using the kernel function without ever explicitly knowing the mapping, e.g., $\theta(u)\theta(v)\equiv K(u,v)$. Some of the more popular known kernels are given below. New kernels are being developed to fit domain specific requirements. 
\begin{center}
	\begin{tabular}{cc}
	$\vartheta(u)$  & $K(u,v)$ \\ \hline 
	Degree $d$ polynomial & $\left(u\cdot v +1\right)^d$ \\ 
	Radial Basis Function & $\exp \left( - \frac{\left\|u-v\right\|^2}{2\sigma}\right) $ \\ 
	Two layer Neural Network& $ sigmoid \left(\eta\left(u\cdot v +1\right)+c\right)$ \\ 
	\end{tabular}
\end{center}
To change from a linear to nonlinear classifier, one must only substitute a kernel evaluation in the objective instead of the original dot product. Thus by changing kernels we can get different highly nonlinear classifiers. No algorithmic changes are required from linear case other than substitute of a kernel evaluation for simple dot product. All the benefits of the original linear SVM methods are maintained. We can train a highly nonlinear classification function such as a polynomial or a radial basis function machine, or a sigmoid neural network using robust, efficient algorithms that have no problem with local minima. By using kernel substitution a linear algorithm (only capable of handling separable data) can be turned into a general nonlinear algorithm. 
%\begin{equation}
%\begin{array}{cc}
%\underline{&}\\  
% & \left(u\cdot v +1\right)^d \\
%Radial Basis Function & \exp \left( - \frac{\left\|u-v\right\|^2}{2\sigma}\right) \\
%Two layer Neural Network & sigmoid \left(\eta\left(u\cdot v +1\right)+c\right) 
% \end{array} 
%\label{eq:diffkernel}
%\end{equation}

%The foundations of Support Vector Machines (SVM) have
\subsubsection{SVM  Classification}
The resulting SVM method can be summarized as follows 
\begin{enumerate}
	\item Selecting the SVM classifier Paramters. Parameter $C$ representing the tradeoff between minimizing the training set error and maximizing the margin. 
	\item Select A kernel function and their kernel parameters. For example for the radial basis function kernel, one must select width of the Gaussian $\gamma$. 
	\item Solve dual QP Equation \ref{eq:dualkernel} or an alternative SVM formulation using an appropriate quadratic programming or linear programming algorithm. 
	\item Get the support vector and recover the emprial threshold  variable $b$. 
	\item Classify new points using $f(x)=sign\left(\sum{i}y_i\alpha_i K\left(x,x_i\right)-b\right)$
\end{enumerate}
The parameters in step 1. are selected using cross validation if sufficient data are available. %However, recent model  selectin strategies can 
\subsubsection {Multi Class Classifiers}

 There is mainly two popular methods to change the binary SVM classification problem into Multiclass classification problem. The first methods is One Versus All (OVA) Strategy where M binary classifiers is constructed one classifier for each class.  The $ith$ classifier is trained using positive example form $ith$ class and negative example from all other classes. The output function $p_i$ represents the distance and classification of the example from the $ith$ classifier. For a new example $x$, OVA SVM strategy assigns it to the class with the largest value of $p_i$.
